Re-Allocation of Resources during Releases \titlenote{Copyright note}

### 2013-10-26 paper1/paper1.tex
updating paper

\title{Re-Allocation of Resources during Releases \titlenote{Copyright information will go here.}}
\numberofauthors{2}
\author{
% 1st. author
\alignauthor
Md Tajmilur Rahman\\
       \affaddr{Concordia University}\\
       \affaddr{Montreal, QC H3G 1M8}\\
       \affaddr{438-932-2288, +1}\\
       \email{rupak.karmadhar@gmail.com}
% 2nd. author
\alignauthor
Peter C. Rigby\\
       \affaddr{Concordia University}\\
       \affaddr{Montreal, QC H3G 1M8}\\
       \affaddr{514-848-2424, +1}\\
       \email{peter.rigby@concordia.ca}
}
\date{20 Oct 2013}
\maketitle
\begin{abstract}
This section will be written at the end.
\end{abstract}
\category{K.6.3}{Software Management}{Software Development; Software Resource Management, Resource Reallocation}
\terms{Experiment, Human Factors, Resource Management, Reallocation}
\keywords{Resource Reallocation, Software Releases} % NOT required for Proceedings
\section{Introduction}
Software projects are notorious for going over budget and schedule. Rush periods are often get seen before a major release that turn the developers into dinosaurs as Frederick Brooks likens in his benchmark study "The Mythical Man Month" \cite{1_brooks}. This "Rush To Release (RTR)" can be prompted either by external forces such as decisions by management to include new features in the release or to release earlier to beat a competitor. Alternatively, the rush may simply be due to inappropriate or unrealistic scheduling. Whatever the reason is it is an obvious. Regardless of the causes, the rush to release stresses developers and often requires developers to work on unusual, high priority or critical areas of the system. In this paper we study how RTR effects project organization and introduces technical debt. The key research questions that we expect to answer with our methodology are as follows:
\begin{enumerate}
\item Do developers work on different areas of the system around the time of release?
\item Are there certain areas of the system that receive increased attention (i.e.\ do developers focus on a smaller set of files around releases)?
\item Do the areas of code that are modified around the time of release have higher defect densities than code that is modified during normal development?
\end{enumerate}

@@

\subsection{Extracting Data}
We went for the VCS of a target project and either mirror the repository or download each revision and commit log history data individually. From VCSs such as CVS we extract the revisions and release information. We then put them into a database. We have used here PSQL database to create tables in, to store our extracted data. These extracted will be analyzed by us later on. Per each revision the information extracted includes the commit id, tree id, author of the revision, date of revision, the name of the revised file, parent and child info for the revision and the detail log information. Once extraction is complete we are ready to partition the version numbers (Section 4.2) and duration  of each release (Section 4.3).

\subsection{Partioning Release Numbers}
We stored the git commit log extracted data into a table named git commit and git revision where all the basic and log information for a particular commit was mentioned in the first table and second one containing which commit belongs to which version of Linux kernel development and change details like path modified, new path created due to the change, how many addition and how many deletion occurred in a particular commit etc. By joining these tables we can easily get the dates of each version and from the version number which is a combination of different types of releases we can get determine which commit belongs to which version and release, and also what type of release that is as well. Figure 1 shows the picture of the data how it looks like.
\begin{figure}
\centering
\psfig{file=versions.png, height=1in, width=3in,}
\caption{Releases extracted from Version}
\end{figure}

@@

\subsection{Get Time-Span Between Release}
Another information that we require is what are the durations of the releases of Linux kernel development. To achieve that we joined the table shown in figure 1 and git\_commit. The following psql query did the job pretty easily for us:
\begin{lstlisting}
select
grt.path as release, grt.major, grt.minor,
grt.micro, grt.rc, gc.committer_dt
from
git_refs_tags grt, git_commit gc
where
gc.commit = grt.commit
order by
gc.committer_dt
\end{lstlisting}
This query gives us all the dates of commit for each and every release, so we can easily find out the duration between two consecutive releases. This information is more important to us because we need to see what is the actual development period and what is the merging or other fixing period within a release time-span and also we need to understand how developers are working, what is the impact of their changes made in the code-base during a particular release period. Table 1 shows a small portion of this information. Preior to calculate the areas of the developers in secion 4.4 we require this information.

\begin{table}[ht]
\caption{Nonlinear Model Results}	 % title of Table
\centering 						% used for centering table
\begin{tabular}{c c c c}				% centered columns (4 columns)
\hline\hline						%inserts double horizontal lines
% table heading
Release & Type & Start Date & End Date \\ [0.5ex]
\hline 							% single horizontal line
% inserting body of the table
linuxv2.6.12-rc2 & rc       & 2005-04-16 & 2005-04-16 \\
linuxv2.6.12-rc3 & rc       & 2005-04-16 & 2005-04-20 \\
linuxv2.6.12-rc4 & rc       & 2005-04-20 & 2005-05-07 \\
linuxv2.6.12-rc5 & rc       & 2005-05-07 & 2005-05-24 \\
linuxv2.6.12-rc6 & rc       & 2005-05-24 & 2005-06-06 \\
linuxv2.6.12        & micro & 2005-06-06 & 2005-06-17 \\
linuxv2.6.13-rc1 & rc       & 2005-06-17 & 2005-06-29 \\
linuxv2.6.13-rc2 & rc       & 2005-06-29 & 2005-07-05 \\
linuxv2.6.13-rc3 & rc       & 2005-07-05 & 2005-07-13 \\
linuxv2.6.13-rc4 & rc       & 2005-07-13 & 2005-07-28 \\
linuxv2.6.13-rc5 & rc       & 2005-07-28 & 2005-08-02 \\
linuxv2.6.13-rc6 & rc       & 2005-08-02 & 2005-08-07 \\
linuxv2.6.13-rc7 & rc       & 2005-08-07 & 2005-08-23 \\
linuxv2.6.13        & micro & 2005-08-23 & 2005-08-28 \\
linuxv2.6.14-rc1 & rc       & 2005-08-28 & 2005-09-12 \\
linuxv2.6.14-rc2 & rc       & 2005-09-12 & 2005-09-19 \\
linuxv2.6.14-rc3 & rc       & 2005-09-19 & 2005-09-30 \\
linuxv2.6.14-rc4 & rc       & 2005-09-30 & 2005-10-10 \\
linuxv2.6.14-rc5 & rc       & 2005-10-10 & 2005-10-20 \\
[1ex]							% adds vertical space
\hline 							% inserts single line
\end{tabular}
\label{table:nonlin} 				% is used to refer this table in the text
\end{table}

\subsection{Calculate Developers' Area}
Developers are working throughout the release time. Developers' are means the files developers work in. But it doesn't refer the to file itself but to understand which files are being touched by which developers. We see that authors are committing files almost everyday throughout a release. We have the commit log data in our hand and we know what is the duration of each releases (i.e.\ start and end dates). Now it is possible to find out which commits were made for which files by a developer within a given time range of a release.

We select all the release information from the table where we stored the releases with their corresponding time-span. After that we select author name and other relevant information from the joining of ``git\_revision'' and ``git\_commits'' tables for every release. We insert these information into a table named ``dev\_area\_rel'' so that we can use the data for our further analysis.

The following pseudo code may show the path to achieve the desired result in this section:

@@

\WHILE{$prev\_release\_date \leftarrow release\_start\_date and release \leftarrow release\_version$}
\IF{$release$}
	\IF{$commit\_date \geq  release\_start\_date$ and $commit\_date < release\_end\_date$}
		\STATE $select author, path, commits, churn$
		\STATE $insert data into table dev\_area\_rel$
	\ENDIF
\ENDIF
\ENDWHILE
\end{algorithmic}
\end{algorithm}

>>>>

\title{Re-Allocation of Resources during Releases \titlenote{Copyright note}}
\numberofauthors{2}
\author{
% 1st. author
\alignauthor
Md Tajmilur Rahman\\
       \affaddr{Concordia University}\\
       \affaddr{Montreal, QC H3G 1M8}\\
       \affaddr{438-932-2288, +1}\\
       \email{mdt\_rahm@encs.concordia.ca}
% 2nd. author
\alignauthor
Peter C. Rigby\\
       \affaddr{Concordia University}\\
       \affaddr{Montreal, QC H3G 1M8}\\
       \affaddr{514-848-2424, +1}\\
       \email{peter.rigby@concordia.ca}
}
\date{20 Oct 2013}
\maketitle
\begin{abstract}
This section will be written at the end.
\end{abstract}
\category{K.6.3}{Software Management}{Software Development; Software Resource Management, Resource Reallocation}
\terms{Experiment, Human Factors, Resource Management, Reallocation}
\keywords{Resource Reallocation, Software Releases} % NOT required for Proceedings
\section{Introduction}
Software projects are notorious for going over budget and schedule. Rush periods are often get seen before a major release that turn the developers into dinosaurs as Frederick Brooks likens in his benchmark study "The Mythical Man Month" \cite{1_brooks}. This "Rush To Release (RTR)" can be prompted either by external forces such as decisions by management to include new features in the release or to release earlier to beat a competitor. Alternatively, the rush may simply be due to inappropriate or unrealistic scheduling. Whatever the reason is it is an obvious. Regardless of the causes, the rush to release stresses developers and often requires developers to work on unusual, high priority or critical areas of the system. In this paper we study how RTR effects project organization and introduces technical debt. The key research questions that we expect to answer with our methodology are as follows:
\begin{enumerate}
\item Do developers work on different areas of the system around the time of release?
\item Are there certain areas of the system that receive increased attention (i.e.\ do developers focus on a smaller set of files around releases)?
\item Do the areas of code that are modified around the time of release have higher defect densities than code that is modified during normal development?
\end{enumerate}

@@

\subsection{Extracting Data}
We went for the VCS of a target project and either mirror the repository or download every revision and commit log history data. From DVCSs such as Git we extract the revisions and release information. We wrote Perl scripts to extract date that was further processed to obtain details. Manual inspection was used to resolve problems and things like that in cases where all automated techniques failed.
We then put them into a database. We have used PSQL database to create tables in, to store our extracted data. These extracted will be analyzed by us later on. Per each revision the information extracted includes the commit id, tree id, author of the revision, date of revision, the name of the revised file, parent and child info for the revision and the detail log information. Once extraction is complete we are ready to partition the version numbers (Section 4.2) and duration  of each release (Section 4.3).

\subsection{Partitioning Release Numbers}
We stored the git commit log extracted data into a table named git commit and git revision where all the basic and log information for a particular commit was mentioned in the first table and second one containing which commit belongs to which version of Linux kernel development and change details like path modified, new path created due to the change, how many addition and how many deletion occurred in a particular commit etc. By joining these tables we can easily get the dates of each version and from the version number which is a combination of different types of releases we can get determine which commit belongs to which version and release, and also what type of release that is as well. Figure 1 shows the picture of the data how it looks like.
\begin{figure}
\centering
\psfig{file=versions.png, height=1in, width=3in,}
\caption{Releases extracted from Version}
\end{figure}

@@

\subsection{Get Time-Span Between Release}
Another information that we require is what are the durations of the releases of Linux kernel development. To achieve that we joined the table in figure 1 and the table ``git\_commit''. The following psql did the job pretty easily for us:
\begin{lstlisting}
select
grt.path as release, grt.major, grt.minor,
grt.micro, grt.rc, gc.committer_dt
from
git_refs_tags grt, git_commit gc
where
gc.commit = grt.commit
order by
gc.committer_dt
\end{lstlisting}
This query gives us all the dates of commit for each and every release, so we can easily find out the duration between two consecutive releases (all including release candidates RC). This information is more important to us because we need to see what is the actual development period and what is the merging or other fixing period within a release time-span and also we need to understand how developers are working, what is the impact of their changes made in the code-base during a particular release period. Table 1 shows a small portion of this information. Preior to calculate the areas of the developers in secion 4.4 we require this information.

\begin{table}[ht]
\caption{Different Releases}  % title of Table
\centering 						% used for centering table
\begin{tabular}{c c c c}				% centered columns (4 columns)
\hline\hline						%inserts double horizontal lines
% table heading
Release & Type & Start Date & End Date \\ [0.5ex]
\hline 							% single horizontal line
% inserting body of the table
linuxv2.6.12-rc2 & rc       & - & 2005-04-16 \\
linuxv2.6.12-rc3 & rc       & 2005-04-16 & 2005-04-20 \\
linuxv2.6.12-rc4 & rc       & 2005-04-20 & 2005-05-07 \\
linuxv2.6.12-rc5 & rc       & 2005-05-07 & 2005-05-24 \\
linuxv2.6.12-rc6 & rc       & 2005-05-24 & 2005-06-06 \\
linuxv2.6.12        & micro & 2005-06-06 & 2005-06-17 \\
linuxv2.6.13-rc1 & rc       & 2005-06-17 & 2005-06-29 \\
linuxv2.6.13-rc2 & rc       & 2005-06-29 & 2005-07-05 \\
linuxv2.6.13-rc3 & rc       & 2005-07-05 & 2005-07-13 \\
linuxv2.6.13-rc4 & rc       & 2005-07-13 & 2005-07-28 \\
linuxv2.6.13-rc5 & rc       & 2005-07-28 & 2005-08-02 \\
linuxv2.6.13-rc6 & rc       & 2005-08-02 & 2005-08-07 \\
linuxv2.6.13-rc7 & rc       & 2005-08-07 & 2005-08-23 \\
linuxv2.6.13        & micro & 2005-08-23 & 2005-08-28 \\
linuxv2.6.14-rc1 & rc       & 2005-08-28 & 2005-09-12 \\
linuxv2.6.14-rc2 & rc       & 2005-09-12 & 2005-09-19 \\
linuxv2.6.14-rc3 & rc       & 2005-09-19 & 2005-09-30 \\
linuxv2.6.14-rc4 & rc       & 2005-09-30 & 2005-10-10 \\
linuxv2.6.14-rc5 & rc       & 2005-10-10 & 2005-10-20 \\
[1ex]							% adds vertical space
\hline 							% inserts single line
\end{tabular}
\label{table:nonlin} 				% is used to refer this table in the text
\end{table}

Note that the very first row has no information for the ``Start Date'' because we don't have any data prior to release ``linuxv2.6.12''. From the extraction we found complete records for releases ``linuxv2.6.13'' to ``linuxv2.6.39'' then ``linuxv3.0'' to ``linuxv3.11''. We have last couple of release candidates and the release date for ``linuxv2.6.12'' and first couple of release candidates for ``linuxv3.12'' which is not useful for our calculation. So we are having 340 releases including major, minor, micro and rc; 40 complete releases from ``linuxv2.6.12'' to ``linuxv3.11''. We have total 400441 distinct commits and for 372943 of them have change information for 77082 different files in our hand where we have 14599 developers' working history among 14621 distinct developers.

\subsection{Calculate Developers' Area}
Developers are working throughout the release time. Developers' Are (DA) means the files developers work in. It doesn't refer the to file itself but it is important to understand which files are being touched by which developers. We see that authors are committing files almost everyday throughout a release. We have the commit log data in our hand and we know what is the duration for each and every releases (i.e.\ start and end dates). Now it is possible to find out which commits were made for which files by a developer within a given time range of a release.

\subsubsection{DA in a Release Period}
We select all the release information from the table where we stored the releases with their corresponding time-span. After that we select author name and other relevant information from the joining of ``git\_revision'' and ``git\_commits'' tables for every release. We insert these information into another table so that we can use the data for our further analysis.

The following pseudo code shows us the way of achieving the desired result in this section:

@@

\WHILE{$prev\_release\_date \leftarrow release\_start\_date\ and\ release \leftarrow release\_version$}
\IF{$release$}
	\IF{$commit\_date \geq  release\_start\_date$ and $commit\_date < release\_end\_date$}
		\STATE $select author, path, commits, churn$
		\STATE $insert data into table dev\_area\_rel$
	\ENDIF
\ENDIF
\ENDWHILE
\end{algorithmic}
\end{algorithm}

A relatively straightforward discipline of Linux Kernel Development team is followed with regard to the merging of patches for each release \cite{14_kernel}.  At the beginning of each development cycle, the "merge window" is said to be opened.  At that time, code which is deemed to be sufficiently stable (and which is accepted by the development community) is merged into the mainline kernel. The bulk of changes for a new development cycle (and all of the major changes) will be merged during this time, at a rate approaching 1,000 changes ("patches," or "changesets") per day. Strategically linux starts a new kernel release with the merging and fixing to get their development branch ready to start development for the next release. So there are two main segments in a release period, one is merge window or merge period (MP) and another is development window what we are going to call Release Development Period (RDP).

Similarly we have calculated DA in both MP and RDP as described in the following subsections.

\subsubsection{DA in a Merge Period}
Merge Perid can be determined from the previous release (any of major, minor or micro releases) date to the first RC date. For example if the date of release for ``linuxv2.6.12''  is 2005-06-17 and date of pushing the first RC for the next release ``linuxv2.6.13-rc1'' is 2005-06-29 then this time period of of 12 days is being called the merge window \cite{14_kernel} or MP. Table 2 shows some merging periods of different releases.

\begin{table}[ht]
\caption{Release Merging Periods}  % title of Table
\centering 						% used for centering table
\begin{tabular}{c c c c}				% centered columns (4 columns)
\hline\hline						%inserts double horizontal lines
% table heading
Release 			& Start Date		& RC Date \\ [0.5ex]
\hline 							% single horizontal line
% inserting body of the table
linuxv2.6.13		& 2005-06-17	& 2005-06-29 \\
linuxv2.6.14		& 2005-08-28	& 2005-09-12 \\
linuxv2.6.15		& 2005-10-27	& 2005-11-11 \\
linuxv2.6.16		& 2006-01-02	& 2006-01-17 \\
linuxv2.6.17		& 2006-03-20	& 2006-04-02 \\
linuxv2.6.18		& 2006-06-17	& 2006-07-06 \\
linuxv2.6.19		& 2006-09-19	& 2006-10-04 \\
linuxv2.6.20  		& 2006-11-29	& 2006-12-13 \\
linuxv2.6.21		& 2007-02-04	& 2007-02-20 \\
linuxv2.6.22		& 2007-04-25	& 2007-05-12 \\
linuxv2.6.23		& 2007-07-08	& 2007-07-22 \\
linuxv2.6.24		& 2007-10-09	& 2007-10-23 \\
linuxv2.6.25		& 2008-01-24	& 2008-02-10 \\
[1ex]							% adds vertical space
\hline 							% inserts single line
\end{tabular}
\label{table:nonlin} 				% is used to refer this table in the text
\end{table}

We understand DA or Developers' Area is the files in the code-base, developers are working from the dataset that we have in our hand. We have the commit logs and every record says about who committed the file when did he/she worked on the file. If we run an SQL query then we can easily find out for every author and for every particular file how many times it's been committed by an author and what changes (here, addition + deletion = churn) he/she has been made. So we are calculating the DA and storing the information into a table. A part of that table is shown in table 3.

\begin{table}[ht]
\caption{DA in Merge Period}  % title of Table
\centering 						% used for centering table
\begin{tabular}{c c c c}				% centered columns (4 columns)
\hline\hline						%inserts double horizontal lines
% table heading
Author		& File Path			& Commits		& Churn \\ [0.5ex]
\hline 							% single horizontal line
% inserting body of the table
D. S. Miller	& include/.../pci.h		& 1				& 8 \\
V. Hanquez	& arch/.../cpu.c		& 1				& 14 \\
A. Bunk		& drivers/.../shmem.c	& 1				& 2 \\
J. Juhl		& arch/.../generic.c		& 1				& 3 \\
[1ex]							% adds vertical space
\hline 							% inserts single line
\end{tabular}
\label{table:nonlin} 				% is used to refer this table in the text
\end{table}

In this table we see in a merging period inside a release period D. S. Miller has made change in pci.h file only once wich the churn number 8. In the similar way we calculate DA in RDP and RTR period as well.

\subsubsection{DA in a RD Period and RTR}
As like MP we can extract the RDP from the. RDP here in Linux Kernel development process is being considered as from the release date of RC1 (i.e.\ first release candidate opening the gate for the development of the next release) to the next release (major, minor or micro) date. For example if the date of release for ``linuxv2.6.13-rc1'' after releasing ``linuxv2.6.12''  is 2005-06-29 and date of publishing the next release ``linuxv2.6.13'' is 2005-08-28 then this time period of of 60 days is being called the RDP. Table 4 is showing some development periods of different releases.

\begin{table}[ht]
\caption{Release Development Periods}  % title of Table
\centering 						% used for centering table
\begin{tabular}{c c c c}				% centered columns (4 columns)
\hline\hline						%inserts double horizontal lines
% table heading
Release 			& RD Start Date	& Release Date \\ [0.5ex]
\hline 							% single horizontal line
% inserting body of the table
linuxv2.6.13		& 2005-06-29	& 2005-08-28 \\
linuxv2.6.14		& 2005-09-12	& 2005-10-27 \\
linuxv2.6.15		& 2005-11-11	& 2006-01-02 \\
linuxv2.6.16		& 2006-01-17	& 2006-03-20 \\
linuxv2.6.17		& 2006-04-02	& 2006-06-17 \\
linuxv2.6.18		& 2006-07-06	& 2006-09-19 \\
linuxv2.6.19		& 2006-10-04	& 2006-11-29 \\
linuxv2.6.20  		& 2006-12-13	& 2007-02-04 \\
linuxv2.6.21		& 2007-02-20	& 2007-04-25 \\
linuxv2.6.22		& 2007-05-12	& 2007-07-08 \\
linuxv2.6.23		& 2007-07-22	& 2007-10-09 \\
linuxv2.6.24		& 2007-10-23	& 2008-01-24 \\
linuxv2.6.25		& 2008-02-10	& 2008-04-16 \\
[1ex]							% adds vertical space
\hline 							% inserts single line
\end{tabular}
\label{table:nonlin} 				% is used to refer this table in the text
\end{table}

So along these RDs of each release period developers get involved in the development for the next release. As we mentioned earlier RTR may happen as the release approaches. Keeping this in our aim list we noticed an important slot in these RDs of releases. We see in between two consecutive releases there is a merging period of time, after that development starts by pushing RCs and once RC releases stops and it takes some days of time for the preperation to publish the release. We are calling this last part as RTR Period because we asume RTR may happen in this period of time before release. Table 5 represents some releases with their corresponding RTR (i.e.\ from end date to rc and date of release).

\begin{table}[ht]
\caption{RTR Periods}  % title of Table
\centering 						% used for centering table
\begin{tabular}{c c c c}				% centered columns (4 columns)
\hline\hline						%inserts double horizontal lines
% table heading
Release 			& RC End Date	& Release Date	& Days \\ [0.5ex]
\hline 							% single horizontal line
% inserting body of the table
linuxv2.6.13		& 2005-06-29	& 2005-08-28	& 5 \\
linuxv2.6.14		& 2005-09-12	& 2005-10-27	& 8 \\
linuxv2.6.15		& 2005-11-11	& 2006-01-02	& 9 \\
linuxv2.6.16		& 2006-01-17	& 2006-03-20	& 8 \\
linuxv2.6.17		& 2006-04-02	& 2006-06-17	& 12 \\
linuxv2.6.18		& 2006-07-06	& 2006-09-19	& 7 \\
linuxv2.6.19		& 2006-10-04	& 2006-11-29	& 14 \\
linuxv2.6.20  		& 2006-12-13	& 2007-02-04	& 10 \\
linuxv2.6.21		& 2007-02-20	& 2007-04-25	& 7 \\
linuxv2.6.22		& 2007-05-12	& 2007-07-08	& 8 \\
linuxv2.6.23		& 2007-07-22	& 2007-10-09	& 9 \\
linuxv2.6.24		& 2007-10-23	& 2008-01-24	& 5 \\
linuxv2.6.25		& 2008-02-10	& 2008-04-16	& 8 \\
[1ex]							% adds vertical space
\hline 							% inserts single line
\end{tabular}
\label{table:nonlin} 				% is used to refer this table in the text
\end{table}
